---
title: "First Analyses for Isotopes February '18"
author: "Joan"
date: "2/10/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This R Markdown document evalulates stable isotopes from sugar pine needles collected in August, 2017

## First look at data
```{r}
library(ggplot2)
##read in data
isotopes = read.csv("2018_IsotopeData.csv", header = T)


names(isotopes)[1:15]=tolower(names(isotopes[1:15]))


summary(lm(d13cpermil~year+treatment, data=isotopes))
summary(lm(d15npermil~year+treatment, data=isotopes))


ggplot(isotopes, aes(year, d13cpermil, color=treatment))+
  geom_smooth()+
  geom_point()+
  theme_classic()

ggplot(isotopes, aes(year, d15npermil, color=treatment))+
  geom_smooth()+
  geom_point()+
  theme_classic()


#c data normal? - yes!
qqnorm(isotopes$d13cpermil)
qqline(isotopes$d13cpermil)

#n data normal? - pretty good
qqnorm(isotopes$d15npermil)
qqline(isotopes$d15npermil)

##are the independent variables (year+treatment) multicollinear?
summary(lm(year~treatment, data=isotopes)) ## nope

##DEALING WITH PAIRED, REPEATED MEASURES
##violation=non-independent variables - we sampled the same individual for six years

##since these are repeat measures (one invidual for 6 years) so we use a Two-Way ANOVA with Repeated Measures
##and add an error term, which controls for the between-pair variation
##We can also control for autocorrelation adding in year as an independent variable into the anova


summary(lm(d13cpermil~year+treatment, data=isotopes))
summary(aov(d13cpermil~year+treatment+Error(pair/(year+treatment)), data=isotopes))
summary(aov(d13cpermil~year+treatment, data=isotopes))

summary(lm(d15npermil~year+treatment, data=isotopes))
summary(aov(d15npermil~year+treatment+Error(pair/(year+treatment)), data=isotopes))
summary(aov(d15npermil~year+treatment, data=isotopes))


##boxplots
ggplot(isotopes, aes(x=year, y=d13cpermil, color=treatment))+
  geom_boxplot(position = "dodge")+
  labs(title="Average C13 between infected and uninfected trees", x="Treatment", y="13C (o/oo")+
  theme_bw()

ggplot(isotopes, aes(x=year, y=d15npermil, color=treatment))+
  geom_boxplot(position = "dodge")+
  labs(title="Average N between infected and uninfected trees", x="Treatment", y="N (o/oo) ")+
  theme_bw()

